Presentations

### 2023-02-22 PhD_pages/Paper_1/data_appendix/1_A_prepare_PIP.qmd
Paper 1: preparing presentation

df_pip["Bottom_50_share"] = df_pip[sum_cols].sum(axis=1)

df_pip["P90_P50_ratio"] = df_pip["decile9_thr"]/df_pip["median"]

@@

df_pip = df_pip.rename(columns={
    'decile10_share':'Top_10_share',
    'mld': 'MLD',
    "country": "Entity",
    "year": "Year"
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top_10_share",
  "Bottom_50_share",
  "MLD",
  "welfare_type"
]

>>>>

df_pip["Bottom 50pc share"] = df_pip[sum_cols].sum(axis=1)

df_pip["P90:P50 ratio"] = df_pip["decile9_thr"]/df_pip["median"]

@@

df_pip = df_pip.rename(columns={
    'decile10_share':'Top 10pc share',
    'mld': 'MLD',
    "country": "Entity",
    "year": "Year"
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top 10pc share",
  "Bottom 50pc share",
  "MLD",
  "welfare_type"
]

### 2023-02-22 PhD_pages/Paper_1/data_appendix/1_B_prepare_WID.qmd
Paper 1: preparing presentation

@@

df_wid = df_wid.rename(columns={
    'P90-P100 - share of the top 10%':'Top_10_share',
    'P0-P50 - share of the bottom 50%':'Bottom_50_share',
    'P99-P100 - share of the top 1%':'Top_1_share'
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top_10_share",
  "Bottom_50_share",
  "inc_concept"
]

@@

df_wid.columns = df_wid.columns.map(lambda x: '{}_{}'.format(x[0], x[1]))

>>>>

```{python}
# I should shift over to this file at some point
# fp = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/raw/wid_indices_992j.csv'

@@

df_wid = df_wid.rename(columns={
    'P90-P100 - share of the top 10%':'Top 10pc share',
    'P0-P50 - share of the bottom 50%':'Bottom 50pc share',
    'P99-P100 - share of the top 1%':'Top 1pc share'
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  'Top 1pc share',
  "Top 10pc share",
  "Bottom 50pc share",
  "inc_concept"
]

@@

df_wid.columns = df_wid.columns.map(lambda x: '{} – {}'.format(x[0], x[1]))

### 2023-02-22 PhD_pages/Paper_1/paper1.qmd
Paper 1: Drafting presentation
Paper 1: preparing presentation

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
format: html
---

::: {.callout-note collapse="true" appearance="minimal"}
### Load packages

```{r}
library(tidyverse)
library(knitr)
library(data.table)

```
:::

@@

>>>>

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
format: html
bibliography: references.bib
execute:
  echo: false
  warning: false
---

::: {.callout-note collapse="true" appearance="minimal"}
### Load packages and data

```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
from IPython.display import display, HTML, IFrame, Markdown
import dataframe_image as dfi
import imgkit

from wrangling_and_plotting_functions import *

pd.set_option('display.max_rows', None)

```

```{python}

fp = 'data_appendix/data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data_appendix/data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data_appendix/data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data_appendix/data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data_appendix/data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data_appendix/data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

# This specifies the +/- for the 'stable' category in the tables below
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ]
            }

```

```{python}
def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'

    if len(selected_vars)>1:
        if all_metrics_requirement:
            msg = f'{msg} – Matched countries only'
        else:
            msg = f'{msg} – All available countries'

    return(msg)

```

@@

Many answers for individual countries, or for particular regions. Can Our understanding of inequality trends *globally*, however, remains fragmented unclear.

For many it is a given . Many papers in the social sciences, or policy notes begin 'decades of rising inequality' – without to specify where. Implying ubiquitous.

Does not reflect the conflicting data.

This stands in contrast with the data in PIP – an important data source. Relied on for SDGs.

Confusiong terrain: e.g. where for SDGs they use PIP; but HDR they use WID.

Our view remains highly fragmented. Different global datasets offer starkly differing or even contradictory views

Despite the wealth of data available, our view of inequality around the world remains fragmented and unclear.

These are questions that we need answers to. Winners and loosers of globalization. Policy to tackle inequality. SDGs. Philanthropy.

That's what this paper is going to do. (Outline structure).

### 2023-02-22 presentations.qmd
Paper 1: preparing presentation

### From $1.90 to $2.15: the updated International Poverty Line
20/01/2023

@@

>>>>

### From $1.90 to $2.15: the updated International Poverty Line
20/01/2023
Department for Social Policy, University of Oxford

@@

### Triangulating a global view on inequality from secondary data sources
16/02/2023
Institute for New Economic Thinking, University of Oxford

:::{.callout collapse="true"}
## Abstract

Many of the key questions and debates of our time necessitate a global view of economic inequality. Thanks to the development and continued expansion of key global databases of harmonised distributional data, we are in a better position than ever to take such a global view.

And yet, on a range of basic questions concerning the nature and evolution of inequality around the world, there is surprisingly little consensus or clarity. Available global datasets offer starkly differing or even contradictory views. The variety of methods, concepts, and sources adopted in different global databases makes it difficult to interpret and synthesise the insights they offer. Despite the wealth of data available, our view of inequality around the world remains highly fragmented.

To help address this, I line up results from three key global datasets on inequality: the World Inequality Database, the World Bank’s Poverty and Inequality Platform and the Luxembourg Income Study. In doing so, I try to triangulate a more cohesive and comprehensive global view of inequality – drawing together those stylised facts that are robust to choices of data source and methodology; highlighting those that are contingent on such choices; and helping to unpick the reasons for such discrepancies.

:::

[Interactive slides](PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd)
